Citation Analysis of Computer Science

### 2019-11-11 Scientometrics/comp.tex
updates

\title{Using Citation Data to Study Computer Science%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Relationship With Other Fields}

@@

\begin{abstract}
Computer science and its applications has experienced dramatic growth and diversification over the last twenty years. Towards understanding the nature of these changes, we analyze a cohort of the computer science literature using a historiographic approach. To understand the features of this cohort and the flow of information within its components, we construct article level clusters whose elements are linked by citations or co-citations, and reconcile them with subject categories in the Scopus database.

\keywords{Bibliometrics \and Clustering \and Knowledge Evolution \and Research Evaluation}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{01A85 \and 01A90} %\and more}
\end{abstract}

@@

Cite Alan Porter on expanding universe versus myopic focus. Cite Traag, Waltman, Lariviere papers. Peter Sjögårde and Per Ahlgren, Loet Leydesdorff, Caroline S. Wagner and Lutz Bornmann, Discontinuities in citation relations among journals: self-organized criticality as a model of scientific revolutions and change, Scientometrics, 10.1007/s11192-018-2734-6, 116, 1, (623-644), (2018). Antonio Perianes-Rodriguez and Javier Ruiz-Castillo, A comparison of the Web of Science and publication-level classification systems of science, Journal of Informetrics, 10.1016/j.joi.2016.10.007, 11, 1, (32-45), (2017). Qi Wang and Ludo Waltman, Large-scale analysis of the accuracy of the journal classification systems of Web of Science and Scopus, Journal of Informetrics, 10.1016/j.joi.2016.02.003, 10, 2, (347-364), (2016). Henry Small, Kevin W. Boyack and Richard Klavans, Identifying emerging topics in science and technology, 
Crossref

@@

\emph{Data} The dblp computer science bibliography consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz from  `https://dblp.uni-trier.de/xml/`. Publications were parsed from the xml source file and loaded into a PostgreSQL database. These parsed records were then matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represents citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par \emph{Clustering by Direct Citation} Graclus software was downloaded from \\`http://www.cs.utexas.edu/users/dml/Software/graclus.html'. The \emph{comp} dataset was cast as an undirected graph and stored in a file with a header line indicating the number of nodes and edges. Graclus accepts the number of clusters required as a parameter, so we used 10,15, 20, 25, 30, 40, and 50 as this parameter and  \emph{comp} as input. Based on criteria articulated in the Louvain and Leiden clustering algorithms, we chose 20 clusters as the number of output clusters. \par \emph{Clustering by Co-Citation} To build clusters using co-citations, we slightly modified the approach of Small and Sweeney (1985) as follows. We first calculated the number of citations accumulated across Scopus for all 634,100 articles in \emph{comp}. After discarding those publications that had not been cited at all, we restricted further analysis to the those in the 90th percentile or higher resulting in a dataset of size 50,869. We then identified 897,407 publications in Scopus that cite the 50,869 previously selected papers from \emph{comp}. The cited references in these papers were restricted to those in the set of 50,869 highly cited papers previously identified. For each publication in turn, all possible reference pairs were generated. The frequency of these co-cited pairs was then computed across the dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. A total of 2,208,135 unique co-cited pairs were thus obtained. These data were represented in an edge list where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. Variable level clustering involves iteratively selecting the edges with the highest normalized co-citation value and extracting connected components from the graph. 

@@

  \includegraphics[scale=0.3]{vlc.pdf}

@@

An issue faced is the generation of very large clusters by chaining through low edge weights. To address this, we used Small and Sweeney's variable clustering approach~\cite{small_clustering_1985} with initial parameters of threshold ($t$) = 0.5 (median normalized co-citation frequency), increment $i = 0.1$, and maximum cluster size, $mcs = 100$. Thus, at start, all edges below the median normalized co-citation frequency are dissolved. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 <= t <= 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster. All clusters generated thus, contained  less than 100 nodes.  Using this approach, 6,822 clusters containing 35,485 nodes were generated. Clusters containing only 2 nodes were then discarded bringing the total number of clusters down to 3,851. Pairwise clustering was then performed on these clusters to generate super-clusters. After 9 rounds of pairwise clustering, 12 clusters were generated resulting from agglomeration of the original 3,851 clusters

@@

Overall strategy:

>>>>

\title{Citation Analysis of Computer Science%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Relationship With Other Fields}

@@

\begin{abstract}
Computer science and its applications has experienced dramatic growth and diversification over the last twenty years. Towards understanding the nature of these changes, we analyze a cohort of the computer science literature using a historiographic approach. For further insight on the features of this cohort and the flow of information within its components, we construct article level clusters whose elements are linked by citations or co-citations, and reconcile them with subject categories in the All Science Journal Classification.

\keywords{Bibliometrics \and Clustering \and Research Evaluation \and Computer Science}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{01A85 \and 01A90} %\and more}

\end{abstract}

@@

%
Computer science, and its applications, has experienced dramatic growth and diversification over the last twenty years. As noted in a 2017 US National Academies Report~\cite{nas_2017}, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission". Additionally, the collective influence of the Internet of Things, `big' data, accessible cloud computing, and advances in artificial intelligence have also been postulated as a recent driver for the growth and evolution~\cite{siebel2019_digital}.

In a historical study of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles), Salton and Bergmark described the global structure of computer science as consisting of three main subareas (i) theoretical foundations, e.g., theory of computation (ii) hardware and computer systems, e.g., architecture (iii) software, e.g, programming systems~\cite{salton_citation_1979}.  Related areas discussed were  (a) mathematics of computing, e.g., numerical analysis (b) special software topics, e.g., operating systems (c) data management and database systems (d) methodologies valid for multiple applications, e.g., algebraic manipulation (e) computer applications, e.g., computer graphics (f) nontechnical aspects, e.g., computer education. In presenting their study, Salton and Bergmark made the key point that the scientific literature serves a rich source of information to study the structure and historical development of a field.

The Computing Classification System (CCS), an ontology, published by the Association for Computing Machinery (ACM), last updated in 2012, consists of 13 top-level areas. https://dl.acm.org/ccs/ccs.cfm \emph{need to figure out to put this in BibTeX} that reflect a more current view of the field. However, given its considerable influence, a description of computer science, as it exists today, must also consider interactions with other fields. An easy way to map scientific publications, especially interdisciplinary articles or those from adjacent fields, to the CCS classification does not exist, however. Other classification systems exist such as the NSF\\ 
\emph{https://www.nsf.gov/statistics/nsf13327/pdf/tabb1.pdf}, WoS \\ 
\emph{https://bit.ly/2NBRqaH}, and the All Science Journal Classification (ASJC) developed and maintained by Scopus 
\\ \emph{insert url} are based on journals that rely on assigning journal-level labels to articles. The limitations of journal-based annotations are well known~\emph{wang and waltman; lariviere paper} and efforts have been made to instead build article level clusters using citations or other criteria for assembly \emph{cite Boyack, Small, Waltman, etc,}.

In this article, we attempt to provide new insights into computer science by matching records from two major data sources. (i) the dblp bibliography- a reference bibliography for computer science\emph{cite} (ii) the Scopus bibliography. which covers the physical science, life sciences, social sciences, and health sciences. Using direct citations and co-citations, we construct clusters of articles that are drawn from both journals and conference proceedings  and reconcile them with the All Science Journal Classification (ASJC) developed and maintained by Scopus.

We observe that....[what do we observe?]

@@

\emph{Data} The dblp computer science bibliography consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz from  `https://dblp.uni-trier.de/xml/`. Publications were parsed from the xml source file and loaded into a PostgreSQL database. The Scopus dataset of over 72,000,000 publications was previously parsed into a custom schema as part of the comnstruction of a larger data platform for research evaluation~\cite{Korobskiy_2019}. Records in the dblp dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represents citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par \emph{Clustering by Direct Citation} Graclus software was downloaded from \\`http://www.cs.utexas.edu/users/dml/Software/graclus.html'. The \emph{comp} dataset was cast as an undirected graph and stored in a file with a header line indicating the number of nodes and edges. Graclus accepts the number of clusters required as a parameter, so we used 10,15, 20, 25, 30, 40, and 50 as this parameter and  \emph{comp} as input. Based on criteria articulated in the Louvain and Leiden clustering algorithms, we chose 20 clusters as the number of output clusters. \par \emph{Clustering by Co-Citation} To build clusters using co-citations, we slightly modified the approach of Small and Sweeney (1985) as follows. We first calculated the number of citations accumulated across Scopus for all 634,100 articles in \emph{comp}. After discarding those publications that had not been cited at all, we restricted further analysis to the those in the 90th percentile or higher resulting in a dataset of size 50,869. We then identified 897,407 publications in Scopus that cite the 50,869 previously selected papers from \emph{comp}. The cited references in these papers were restricted to those in the set of 50,869 highly cited papers previously identified. For each publication in turn, all possible reference pairs were generated. The frequency of these co-cited pairs was then computed across the dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. A total of 2,208,135 unique co-cited pairs were thus obtained. These data were represented in an edge list where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. Variable level clustering involves iteratively selecting the edges with the highest normalized co-citation value and extracting connected components from the graph. 

@@

  \includegraphics[scale=0.2]{vlc.pdf}

@@

An issue is the generation of very large clusters by chaining through low edge weights. To address this, we used Small and Sweeney's variable clustering approach~\cite{small_clustering_1985} with initial parameters of threshold ($t$) = 0.5 (median normalized co-citation frequency), increment $i = 0.1$, and maximum cluster size, $mcs = 100$. Thus, at start, all edges below the median normalized co-citation frequency are dissolved. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 <= t <= 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster. All clusters generated thus, contained  less than 100 nodes.  Using this approach, 6,822 clusters containing 35,485 nodes were generated. Clusters containing only 2 nodes were then discarded bringing the total number of clusters down to 3,851. Pairwise clustering was then performed on these clusters to generate super-clusters. After 9 rounds of pairwise clustering, 12 clusters were generated resulting from agglomeration of the original 3,851 clusters

@@

Cite Alan Porter on expanding universe versus myopic focus. Cite Traag, Waltman, Lariviere papers. Peter Sjögårde and Per Ahlgren, Loet Leydesdorff, Caroline S. Wagner and Lutz Bornmann, Discontinuities in citation relations among journals: self-organized criticality as a model of scientific revolutions and change, Scientometrics, 10.1007/s11192-018-2734-6, 116, 1, (623-644), (2018). Antonio Perianes-Rodriguez and Javier Ruiz-Castillo, A comparison of the Web of Science and publication-level classification systems of science, Journal of Informetrics, 10.1016/j.joi.2016.10.007, 11, 1, (32-45), (2017). Qi Wang and Ludo Waltman, Large-scale analysis of the accuracy of the journal classification systems of Web of Science and Scopus, Journal of Informetrics, 10.1016/j.joi.2016.02.003, 10, 2, (347-364), (2016). Henry Small, Kevin W. Boyack and Richard Klavans, Identifying emerging topics in science and technology, 
Crossref
Overall strategy:
